---
title: "Day 1 - Methods"
author: "Klaus Langenheldt"
date: "03/05/2018"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is Going on?
```{r, warning = FALSE, echo = T}
library(forecast)
data = mdeaths # change this line to provide new data
model = stl(data, t.window=15, s.window="periodic", robust=TRUE)
plot(model)
```


## Information on Functions/Parameters
```{r, echo=T}
?stl
```

<https://otexts.org/fpp2/stl.html>


## Data Handling & Visualisation
```{r AirPassengers0, echo = TRUE}
# load data
data = AirPassengers

# basic statistics
summary(data)
```

## Visualisation
```{r AirPassengers, echo = TRUE}
plot(AirPassengers)
```

## Data Inspection
```{r, echo = TRUE}
plot(density(AirPassengers))
```

## Decompositon
```{r, warning = FALSE, echo = TRUE}
library(forecast)
fit <- stl(AirPassengers, t.window=15, s.window="periodic", robust=TRUE)
plot(fit)
```

## Seasonal Analysis
```{r, warning = FALSE, echo = TRUE}
boxplot(AirPassengers~cycle(AirPassengers))
```

# The `forecast` package

## The `forecast` package
- The holy grail of R time series forecasting
- Includes all the traditional models
  - AR
  - MA
  - ARIMA
  - ...
  
## Basic Forecasting Methods
Simple Average
$$\hat{y}_{T+h|T} = \bar{y} = (y_{1}+\dots+y_{T})/T$$
Na誰ve Method
$$\hat{y}_{T+h|T} = y_{T}$$
Seasonal Na誰ve Method
$$\hat{y}_{T+h|T} = y_{T+h-m(k+1)}$$

## Comparing Basic Forecasting Methods
```{r results='hide', message=FALSE, warning=FALSE}
library(fpp)
library(ggplot2)
# Set training data from 1992-2007
beer2 <- window(ausbeer,start=1992,end=c(2007,4))
# Plot some forecasts
autoplot(beer2) +
  autolayer(meanf(beer2, h=11),
    series="Mean", PI=FALSE) +
  autolayer(naive(beer2, h=11),
    series="Na誰ve", PI=FALSE) +
  autolayer(snaive(beer2, h=11),
    series="Seasonal na誰ve", PI=FALSE) +
  ggtitle("Forecasts for quarterly beer production") +
  xlab("Year") + ylab("Megalitres") +
  guides(colour=guide_legend(title="Forecast"))
```


# That's TS Forecasting


## Predict using Moving Average (MA)
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ma(data, order=2)
forecast = forecast(model, h=10)
plot(forecast)
```


## Predict using Moving Average (MA)
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ma(data, order=2)
forecast = forecast(model, h=30)
plot(forecast)
```


# That's also TS Forecasting


## Prediction Intervals
```{r, warning = FALSE, echo = F}
data = ts(AirPassengers) # convert to ts object
model = ma(data, order=2)
forecast = forecast(model, h=30)
plot(forecast)
```

<https://www.otexts.org/fpp2/prediction-intervals.html>


## Predict using Auto Regression (AR)
#### AR(5)
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ar(data, order=5)
forecast = forecast(model, h=30)
plot(forecast)
```

## Predict using Auto Regression (AR)
#### AR(10)
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ar(data, order=10)
forecast = forecast(model, h=30)
plot(forecast)
```

## Predict using Auto Regression (AR)
Higher order, higher pattern recognition (now AR(13))
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ar(data, order=13)
forecast = forecast(model, h=30)
plot(forecast)
```

## Predict using Auto Regression (AR)
Long time-frames: mean-convergence and zero trend
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
model = ar(data, order=20)
forecast = forecast(model, h=365)
plot(forecast)
```

## Predict using auto.ARIMA
$$AR + I + MA$$

or rather

$$AR(x) + I(x) + MA(x)$$
where $x$ is the 'order' (plain english: how often it is applied)

## What is $I$?
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers) # convert to ts object
diff = diff(data, differences = 1) # create I of n-th order
plot(diff)
```

## Predict using auto.ARIMA (4 years)
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers)
model = auto.arima(data)
forecast = forecast(model, h=30)
plot(forecast)
```


## Long-Term Forecast using auto.ARIMA
```{r, warning = FALSE, echo = TRUE}
data = ts(AirPassengers)
model = auto.arima(data)
forecast = forecast(model, h=120)
plot(forecast)
```


# Practical Challenge

## Predict `mdeaths`
- Predict `mdeaths` with `auto.arima` for 3 years
- `mdeaths` is split into months, hence one year equals `h=12`
- The dataset:
```{r, warning = FALSE, echo = TRUE}
data = ts(mdeaths)
```


## Solution
```{r, warning = FALSE, echo = TRUE}
data = ts(mdeaths)
model = auto.arima(data)
forecast = forecast(model, h=36)
plot(forecast)
```

## What to do?
- Change forecast method
- Transform data
- Tune $AR(x) + I(x) + MA(x)$ orders manually

# Recap

## Recap
- Forecasting can be as simple as taking the average
- More complex forecasting follows the same approach: extrapolating from the past
- The `forecast` package is the holy grail of forecasting
- `auto.arima()` provides a way to do automated forecasting with classic methods


# Cutting-edge forecasting with `prophet`

## Traditional Methods
- Not made for large-scale data
- Need a lot of manual inspection and tuning
- Sensitive to outliers and missing data
- Requires an in-house specialist in **each** department
- In short: costly to deploy & to scale

## Modern Methods
- Generalised Additive Models
- Memoryless Neural Networks (NN)
- Recurrent Neural Networks (RNN/LTSM)


## Modern Method: `prophet`
![](https://research.fb.com/wp-content/themes/fb-research/images/branding/FB_logo.svg)

<https://research.fb.com/prophet-forecasting-at-scale/>


## Modern Method: `prophet`
```{r include=FALSE}
source("data/prophet-AirPassengers-short.R")
```
```{r echo=T, eval=F}
m = prophet(df)
future <- make_future_dataframe(m, periods = 12, freq = 'month')
forecast <- predict(m, future)
```

See `data/prophet-AirPassengers-short.R` for full code

## Modern Method: `prophet`
```{r, warning = FALSE, echo = TRUE}
plot(m, forecast)
```

## Modern Method: `prophet`
```{r, warning = FALSE, echo = TRUE}
prophet_plot_components(m, forecast)
```

## How Does `prophet` Work?
- **NOT** AutoRegressive (rather, finds a _curve_)
- Robust against
    - missing data
    - outliers
    - significant trend changes
- Fully automatic
- Fast at internet-scale data

$$Y = level + trend + seasonality + holidays + noise$$

# Practical Challenge

## Practical Challenge
- Open `day1/data/prophet-AirPassengers-long.R`
- Run the file
    - it will predict 10 years using
        - `prophet` and
        - `auto.arima`
- You will receive **two** plots (use the arrows to navigate between them)
- Which forecast is more sensible?

# Recap

## Recap
- Old methods rule the game
- ...but modern methods make way for easier forecasting - they _democratise_ forecasting
- For example, `prophet` offers a low-cost, high-scale platform
- A cost-effcient way to experiment (and to forecast, too)
- you can do **low-cost** experiments with **automated** methods, use e.g.
    - `prophet`
    - `auto.arima()` from the `forecast` package
- Try both!
